---
title: "GA6.qmd"
author: "Anna Kolganova"
format: html
theme: darkly 
editor: visual
date: "2025-11-23"
---

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed).

First, I created a repo:

\`\`\`{r}

```         
git init

git add GA6.qmd

git commit -m "Committing to the main document to set up"
```

R Basics:

1.  Create a character vector `fav_food` with five names of your favorite food. Replace third element of the vector with a bird name and extract first and fifth element.

    ```{r}
    fav_food <- c("asparagus", "apples" , "potatoes", "tofu", "beans")

    fav_food

    fav_food[3] <- "bold eagle"

    fav_food

    fav_food[c(1, 5)]
    ```

2.  

    ```{r}
    library(tidyverse)

    require(tidyverse)

    meta <- read_tsv(file = "garrigos-data/meta/metadata.tsv")

    str(meta)

    'The variables are time and treatment and sample id. there are 22 rows.'

    meta$time("10dpi")
    ```

meta \|\> mutate( time = sub(pattern = "10dpi", replacement = "10_dpi", x = time),

treatment = sub(pattern = "cathemerium", replacement = "catch", x = treatment) )

4.  require(ggplot2)

    ```         
    data(package = "ggplot2")
    > midwest_datasets <- ggplot2::midwest
    ```

midwest_datasets

5.  midwest_datasets \|\>

filter(poptotal \> 30000, popdensity \> 800)

6.  require(dplyr)

new_dataset \<- midwest_datasets \|\>

select(1:11) \|\>

mutate(asian_ameri = popasian / popamerindian) \|\>

arrange(desc(asian_ameri))

new_dataset

7.  Note: I wasn't sure if you wanted the mean calculated for the entire midwest_dataset or for new_dataset, so I did both

mean_pop \<- midwest_datasets \|\>

group_by(state) \|\>

summarise(poptotal = mean(poptotal))

mean_pop \<- new_dataset \|\>

group_by(state) \|\>

summarise(poptotal = mean(poptotal))

git add GA6.qmd

git commit -m "Committing to the main document second time"

require(ggplot2)

8.  ggplot(midwest_datasets, aes(x = state, y = area)) +

    geom_violin() +

    facet_wrap(\~ category) +

    theme_light() +

    labs(title = "Violin plot of area by category and state")

9.  ggplot(midwest_datasets, aes(x = state, y = poptotal)) +

    geom_violin() +

    geom_jitter(aes(color = category), width = 0.08, alpha = 0.5) +

    facet_wrap(\~ category) +

    theme_light() +

    labs(title = "Violin plot of area by category and state")

    This point color is local mapping because we assigned color by category inside geom_jitter, not directly inside ggplot(). It means that n other function is affected other than geom_jitter().

10. ggplot(midwest_datasets, aes(x = state, y = poptotal)) +

    geom_violin() +

    geom_jitter(aes(color = state), width = 0.08, alpha = 0.5) +

    facet_wrap(\~ category) +

    theme_bw() +

    labs(title = "Violin plot of area by category and state")

11. 

\`\`\`{r violin_plot, fig.cap="Violin plot of population by state and category"}

\`\`\`r

state_colors \<- c(

IL = "purple",

IN = "green",

MI = "orange",

OH = "maroon",

WI = "pink"

)

violin_plot \<- ggplot(midwest_datasets, aes(x = state, y = poptotal)) +

geom_violin(fill = "grey90") +

geom_jitter(aes(color = state), width = 0.08, alpha = 0.6) +

scale_color_manual(values = state_colors) +

facet_wrap(\~ category) +

theme_bw()

\`\`\`{r violin_plot, fig.cap="Population by State and Category"}

state_colors \<- c(

IL = "purple",

IN = "green",

MI = "orange",

OH = "brown",

WI = "pink"

)

violin_plot \<- ggplot(midwest_datasets, aes(x = state, y = poptotal)) +

geom_violin(fill = "grey90") +

geom_jitter(aes(color = state), width = 0.08, alpha = 0.6) +

scale_color_manual(values = state_colors) +

facet_wrap(\~ category) +

theme_bw()

\`\`\`
